---
title: "Untitled"
output: html_document
date: "2024-04-24"
---

```{r include=FALSE}
install.packages("tidyverse")
install.packages("tidyr")
install.packages("tzdb")
install.packages("timechange")
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
install.packages("masked")

library(conflicted)

```

```{r reading in, include=FALSE}
setwd("C:/Users/mf767/OneDrive - University of Bath")
roots <- read.csv("Roots.csv", h=T)
leaves <- read.csv("Above_ground.csv", h =T)
exp_2 <- read.csv("Experiment_2.csv", h = T)
MFC_mV <- read.csv("MFC_mV.csv", h=T)
MFC_uA <- read.csv("MFC_uA.csv", h=T)
EV_uA <- read.csv("EV_uA.csv", h=T)
#average till width
leaves$till.avg <- mean(leaves$Till.width.1:leaves$Till.width.3)
leaves <- leaves %>% mutate_at(vars(Till.width.3), as.numeric)
leaves$till.avg <- rowMeans(leaves[,4:6], na.rm = TRUE)
new_cols <- c("treatment", "no.")
leaves <- separate(leaves, col = Cell_ID, into = new_cols, sep = "_" )
chl <- read.csv("Chlorophyll.csv",h=T)
new_cols <- c("treatment", "no.")
chl <- separate(chl, col = Cell_ID, into = new_cols, sep = "_" )

```


```{r shapiro test function dont run}
perform_shapiro_test <- function(data, group_column) {
  # Find numeric columns
  numeric_cols <- sapply(data, is.numeric)
  
  # Extract unique group values
  group_values <- unique(data[[group_column]])
  
  # Initialize an empty list to store results
  shapiro_results <- list()
  
  # Iterate over each numeric column
  for (col_name in names(data)[numeric_cols]) {
    # Iterate over each group
    p_values <- sapply(group_values, function(group_value) {
      # Filter data by group
      group_data <- data[data[[group_column]] == group_value, col_name]
      
      tryCatch(
        {
          # Perform Shapiro-Wilk test
          shapiro_result <- shapiro.test(group_data)
          # Extract p-value
          shapiro_result$p.value
        },
        error = function(e) {
          # If an error occurs, return NA
          NA
        }
      )
    })
    
    # Store p-values for each group in the column
    shapiro_results[[col_name]] <- p_values
  }
  
  return(shapiro_results)
}

```

```{r shapiro test and bf test on leaves-treatment, include=FALSE}
install.packages("onewaytests")
library(onewaytests)
install.packages("car")
library(car)
install.packages("FSA")
library(FSA)
```


```{r shapiro test and bf test on leaves-treatment dont run}
leaves_no_MFC <-leaves %>% filter(treatment  == "EV"|treatment == "EC"| treatment == "C")

perform_shapiro_test(leaves_no_MFC,"treatment")
leveneTest(dry_mass ~ treatment, data = leaves_no_MFC)
leveneTest(Wet_Mass ~ treatment, data = leaves_no_MFC)
leveneTest(till.avg ~ treatment, data = leaves_no_MFC)



bf.test(dry_mass ~ treatment, data = leaves_no_MFC)
bf.test(Wet_Mass ~ treatment, data = leaves_no_MFC)
bf.test(till.avg ~ treatment, data = leaves_no_MFC)


# Fit your model (replace 'model' with your actual model)
model <- lm(y ~ x1 + x2, data = )

# Extract residuals from the model
residuals <- resid(model)

# Create a Q-Q plot
qqnorm(residuals)
qqline(residuals)

data(mtcars)


```


```{r stats for leaves-treatment}

failed_MFC <- c(14,16)
leaves <-leaves[-failed_MFC, ]
failed_MFC_chl <- c(2,4)
chl <- chl[-failed_MFC_chl, ]

#wet mass
kruskal.test(Wet_Mass ~ treatment, data = leaves)

dunnTest(Wet_Mass ~ treatment,
  data = leaves,
  method = "holm"
)
#dry mass
kruskal.test(dry_mass ~ treatment, data = leaves)
dunnTest(dry_mass ~ treatment,
  data = leaves,
  method = "holm"
)
#till avg
kruskal.test(till.avg ~ treatment, data = leaves)

dunnTest(till.avg ~ treatment,
  data = leaves,
  method = "holm"
)

#chl a
kruskal.test(Chl_a_mg ~ treatment, data = chl)

dunnTest(Chl_a_mg ~ treatment,
  data = chl,
  method = "holm"
)
#chl b
kruskal.test(Chl_b_mg ~ treatment, data = chl)

dunnTest(Chl_b_mg ~ treatment,
  data = chl,
  method = "holm"
)

#total chl (a+b)
kruskal.test(total_chl ~ treatment, data = chl)

dunnTest(total_chl ~ treatment,
  data = chl,
  method = "holm"
)


```


```{r data summary function}
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      max = max(x[[col]], na.rm=TRUE),
      min = min(x[[col]], na.rm=TRUE),
      sd = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```


```{r boxplots for leaves-treatment}
summary_leaves_wet <- data_summary(leaves, varname = "Wet_Mass", groupnames = c("treatment"))
summary_leaves_dry <- data_summary(leaves, varname = "dry_mass", groupnames = c("treatment"))
summary_leaves_till.avg <- data_summary(leaves, varname = "till.avg", groupnames = c("treatment"))
summary_chl_a <- data_summary(chl, varname = "Chl_a_mg", groupnames = c("treatment"))
summary_chl_b <- data_summary(chl, varname = "Chl_b_mg", groupnames = c("treatment"))
summary_chl_total <- data_summary(chl, varname = "total_chl", groupnames = c("treatment"))




library(ggplot2)
```


```{r wet mass boxplot leaves}
Wet_Mass_box <- ggplot(leaves, aes(x=treatment, y=Wet_Mass , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Green Wet Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_leaves_wet, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))


```


```{r dry mass boxplot leaves}
Dry_Mass_box <- ggplot(leaves, aes(x=treatment, y=dry_mass , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Green Dry Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_leaves_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r till avg boxplot leaves}
Till_avg_box <- ggplot(leaves, aes(x=treatment, y=till.avg , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Average Till Width") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_leaves_till.avg, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r chla  boxplot leaves}
chl_a_box <- ggplot(chl, aes(x=treatment, y=Chl_a_mg , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Chl a") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_chl_a, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r chlb  boxplot leaves}
chl_b_box <- ggplot(chl, aes(x=treatment, y=Chl_b_mg , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Chl b") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_chl_b, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r chl total  boxplot leaves}
chl_total_box <- ggplot(chl, aes(x=treatment, y=total_chl , fill = treatment)) + 
   labs(x= "Cell Treatment", y= "Chl total") + 
scale_fill_discrete(name = "Legend", labels = c("Control", "Electrode Control", "External Voltage"))+
   geom_errorbar(data = summary_chl_total, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```

```{r Roots removing MFC}
new_cols <- c("treatment", "no.", "LorR")
root_ID <- roots$Cell_ID
roots <- separate(roots, col = Cell_ID, into = new_cols, sep = "_" )
roots$root_ID <- root_ID

roots$treatment_LorR <- paste0(roots$treatment, "_", roots$LorR)
roots$treatment_no <- paste0(roots$treatment,"_",roots$no.)

roots_no_MFC <-roots %>% filter(treatment  == "EV"|treatment == "EC"| treatment == "C")

failed_MFC_roots <- c(27,28,31,32)
roots_1 <-roots[-failed_MFC_roots,]

```


```{r dry roots box plot}
Dry_Mass_summary_1 <- data_summary(roots_1, varname = "Dry_Mass", groupnames = c("treatment_LorR"))
new_cols <- c("treatment", "LorR")
Dry_Mass_summary_1 <- separate(Dry_Mass_summary_1, col = treatment_LorR, into = new_cols, sep = "_" )


dry_box_root <- ggplot(roots_1, aes(x=treatment, y=Dry_Mass , fill = LorR)) + 
  labs(x= "Cell Treatment", y= "Root mass") + 
   geom_errorbar(data = Dry_Mass_summary_1, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(0.8),stat = "identity")+
  scale_fill_discrete(name = "Legend", labels = c("C_L", "C_R", "EC_L", "EC_R", "MFC_L", "MFC_R", "EV_L", "EV_R"))+
  geom_boxplot(position = position_dodge(0.8)) +
  scale_x_discrete(limits = c("C", "EC", "MFC", "EV"))+
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
colours <- c("darkgreen", "lightgreen", "darkgreen", "lightgreen", "darkgreen", "lightgreen", "darkgreen", "lightgreen")
dry_box_root + scale_fill_manual(values = colours) 
```


```{r wet roots box plot}
Wet_Mass_summary_1 <- data_summary(roots_1, varname = "Wet_Mass", groupnames = c("treatment_LorR"))
new_cols <- c("treatment", "LorR")
Wet_Mass_summary_1 <- separate(Wet_Mass_summary_1, col = treatment_LorR, into = new_cols, sep = "_" )



wet_box_root <- ggplot(roots_1, aes(x=treatment, y=Wet_Mass , fill = LorR)) + 
  labs(x= "Cell Treatment", y= "Root mass") + 
   geom_errorbar(data = Wet_Mass_summary_1, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(0.8),stat = "identity")+
  scale_fill_discrete(name = "Legend", labels = c("C_L", "C_R", "EC_L", "EC_R", "MFC_L", "MFC_R", "EV_L", "EV_R"))+
  geom_boxplot(position = position_dodge(0.8)) +
  scale_x_discrete(limits = c("C", "EC", "MFC", "EV"))+
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
colours <- c("darkgreen", "lightgreen", "darkgreen", "lightgreen", "darkgreen", "lightgreen", "darkgreen", "lightgreen")
wet_box_root + scale_fill_manual(values = colours) 
```

```{r creating sub dataset for each treatment roots}
roots_no_MFC$LorR <- as.factor(roots_no_MFC$LorR)
roots_1$LorR <- as.factor(roots_1$LorR)
EV<-roots_no_MFC %>% filter(treatment  == "EV")
EC<-roots_no_MFC %>% filter(treatment  == "EC")
C<-roots_no_MFC %>% filter(treatment  == "C")
MFC<-roots_1 %>% filter(treatment  == "MFC")
```


```{r boxplots for EV-LorR dont run}


EV_roots_wet_mass <- ggplot(EV, aes(x=LorR, y=Wet_Mass , fill = LorR)) + 
  geom_boxplot() + labs( x= "Left vs Right", y= "Wet Mass") + scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))

EV_summary_wet <- data_summary(EV, varname = "Wet_Mass", groupnames = c("LorR"))
EV_summary_dry <- data_summary(EV, varname = "Dry_Mass", groupnames = c("LorR"))

EV_roots_wet_mass <- ggplot(EV, aes(x=LorR, y=Wet_Mass , fill = LorR)) + 
   labs(x= "Left vs Right", y= "Wet Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))+
   geom_errorbar(data = EV_summary_wet, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))

EV_roots_dry_mass <- ggplot(EV, aes(x=LorR, y=Dry_Mass , fill = LorR)) + 
   labs(x= "Left vs Right", y= "Dry Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))+
   geom_errorbar(data = EV_summary_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```



```{r boxplots for MFC-LorR dont run}
MFC<-roots_1 %>% filter(treatment  == "MFC")

MFC_roots_wet_mass <- ggplot(MFC, aes(x=LorR, y=Wet_Mass , fill = LorR)) + 
  geom_boxplot() + labs( x= "Left vs Right", y= "Wet Mass") + scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))

MFC_summary_wet <- data_summary(MFC, varname = "Wet_Mass", groupnames = c("LorR"))
MFC_summary_dry <- data_summary(MFC, varname = "Dry_Mass", groupnames = c("LorR"))

MFC_roots_wet_mass <- ggplot(MFC, aes(x=LorR, y=Wet_Mass , fill = LorR)) + 
   labs(x= "Left vs Right", y= "Wet Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))+
   geom_errorbar(data = MFC_summary_wet, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))

MFC_roots_dry_mass <- ggplot(MFC, aes(x=LorR, y=Dry_Mass , fill = LorR)) + 
   labs(x= "Left vs Right", y= "Dry Mass") + 
scale_fill_discrete(name = "Legend", labels = c("Left - No electrodes", "Right- Electrodes"))+
   geom_errorbar(data = MFC_summary_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```




```{r wilcox for EV-LorR}
wilcox.test(Dry_Mass ~ LorR, data = EV,
        var.equal = TRUE, alternative = "less")
wilcox.test(Dry_Mass ~ LorR, data = EV,
        var.equal = TRUE, alternative = "less",paired = FALSE)
wilcox.test(Wet_Mass ~ LorR, data = EV,
        var.equal = TRUE, alternative = "less")
```


```{r wilcox for EC-LorR}
wilcox.test(Dry_Mass ~ LorR, data = EC,
        var.equal = TRUE, alternative = "less")


wilcox.test(Wet_Mass ~ LorR, data = EC,
        var.equal = TRUE, alternative = "less")
```


```{r wilcox for C-LorR}
wilcox.test(Dry_Mass ~ LorR, data = C,
        var.equal = TRUE, alternative = "less")


wilcox.test(Wet_Mass ~ LorR, data = C,
        var.equal = TRUE, alternative = "less")
```


```{r wilcox for MFC-LorR}
wilcox.test(Dry_Mass ~ LorR, data = MFC,
        var.equal = TRUE, alternative = "less")


wilcox.test(Wet_Mass ~ LorR, data = MFC,
        var.equal = TRUE, alternative = "less")

```




```{r EV_uA dont run}
get_max_values <- function(df, Cell_ID_col){
  max <- aggregate(. ~ get(Cell_ID_col), data = df, FUN = max)
  return(max)
}

EV_uA_mean <- c("11.93333333",	"14.68333333",	"5.166666667",	"7.583333333")

EV_uA_Max <-get_max_values(EV_uA, "Cell_ID")

EV_above_ground <- leaves %>% filter(treatment  == "EV")
plot(EV_uA_Max$uA , EV_above_ground$Wet_Mass)
uA_EV <- as.data.frame(EV_uA_Max$uA)
uA_EV$Cell_ID <- EV_uA_Max$`get(Cell_ID_col)`
uA_EV$Wet_Mass <- EV_above_ground$Wet_Mass
uA_EV$Dry_Mass <- EV_above_ground$dry_mass
uA_EV$till.avg <- EV_above_ground$till.avg
uA_EV$Mean_uA <- c(EV_uA_mean)

colnames(uA_EV) <- c("Max_uA", "Cell_ID", "Wet_Mass", "Dry_Mass", "Till.avg", "Mean_uA")

EV_Roots_R <- EV %>% filter(LorR  == "R")

uA_EV$Dry_Mass_Roots_R <- EV_Roots_R$Dry_Mass
uA_EV$Wet_Mass_Roots_R <- EV_Roots_R$Wet_Mass

uA_EV$Max_uA <- as.numeric(uA_EV$Max_uA)
uA_EV$Mean_uA <- as.numeric(uA_EV$Mean_uA)

```





```{r boxplot EV_uA - EV_Roots_R dont run}

summary_uA_EV_dry <- data_summary(uA_EV, varname = "Dry_Mass",groupnames = c("Cell_ID"))
summary_uA_EV_wet <- data_summary(uA_EV, varname = "Wet_Mass",groupnames = c("Cell_ID"))
summary_uA_EV <- data_summary(EV_uA, varname = "uA",groupnames = c("Cell_ID"))

ggplot(EV_uA, aes(x=Cell_ID, y=uA, fill = Cell_ID) )+
  geom_errorbar(data = summary_uA_EV, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot()+
  geom_point(data= uA_EV, aes(y = Dry_Mass_Roots_R*10 , x = Cell_ID), color = "red", shape = 19, size = 4)+
  labs(y = "uA",x = "Cell_ID" )+
  theme(axis.title.y = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black"))+
  scale_y_continuous(sec.axis = sec_axis(~./10, name = "Dry Mass"))+
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))

ggplot(EV_uA, aes(x=Cell_ID, y=uA, fill = Cell_ID) )+
  geom_errorbar(data = summary_uA_EV, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot()+
  geom_point(data= uA_EV, aes(y = Wet_Mass_Roots_R , x = Cell_ID), color = "red", shape = 19, size = 4)+
  labs(y = "uA",x = "Cell_ID" )+
  theme(axis.title.y = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black"))+
  scale_y_continuous(sec.axis = sec_axis(~., name = "Wet Mass"))+
    theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))


```

```{r cor test EV_uA - Mass dont run}
cor_test <- cor.test(uA_EV$Max_uA, uA_EV$Wet_Mass, method = "pearson")
cor_test

cor_test <- cor.test(uA_EV$Max_uA, uA_EV$Dry_Mass, method = "pearson")
cor_test


cor_test <- cor.test(uA_EV$Mean_uA, uA_EV$Dry_Mass, method = "pearson")
cor_test

cor_test <- cor.test(uA_EV$Mean_uA, uA_EV$Wet_Mass, method = "pearson")
cor_test
```


```{r experiment 2 setup}
new_cols <- c("soil", "no.")
exp_2$ID <- exp_2$ID
exp_2 <- separate(exp_2, col = ID, into = new_cols, sep = "_" )
```


```{r experiment 2 wilcox}
exp_2$soil <- as.factor(exp_2$soil)
levels(exp_2$soil)
wilcox.test(Wet_Mass ~ soil, data = exp_2,alternative = "less" )
wilcox.test(Root_Length ~ soil, data = exp_2,alternative = "less" )
wilcox.test(Height ~ soil, data = exp_2,alternative = "less" )
wilcox.test(total_dry_mass ~ soil, data = exp_2,alternative = "less" )
wilcox.test(roots_dry_mass ~ soil, data = exp_2,alternative = "less" )
wilcox.test(aerial_dry_mass ~ soil, data = exp_2,alternative = "less" )




```

```{r total wet mass boxplot experiment 2}
summary_N_Wet <- data_summary(data = exp_2,varname = "Wet_Mass",groupnames = c("soil"))
summary_N_Height <- data_summary(data = exp_2,varname = "Height",groupnames = c("soil"))
summary_N_Root_len <- data_summary(data = exp_2,varname = "Root_Length",groupnames = c("soil"))
summary_N_total_dry <- data_summary(data = exp_2,varname = "total_dry_mass",groupnames = c("soil"))
summary_N_Root_dry <- data_summary(data = exp_2,varname = "roots_dry_mass",groupnames = c("soil"))
summary_N_aerial_dry <- data_summary(data = exp_2,varname = "aerial_dry_mass",groupnames = c("soil"))







exp2_mass_wet <- ggplot(exp_2, aes(x=soil, y=Wet_Mass , fill = soil)) + 
   geom_errorbar(data = summary_N_Wet, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "Wet Mass") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r height boxplot experiment 2}
exp2_height <- ggplot(exp_2, aes(x=soil, y=Height , fill = soil)) + 
     geom_errorbar(data = summary_N_Height, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "Height") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r root length boxplot experiment 2}
exp2_root_length <- ggplot(exp_2, aes(x=soil, y=Root_Length , fill = soil)) + 
     geom_errorbar(data = summary_N_Root_len, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "Root Length") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r total dry mass boxplot experiment 2}
exp2_total_dry <- ggplot(exp_2, aes(x=soil, y=total_dry_mass , fill = soil)) + 
     geom_errorbar(data = summary_N_total_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "total_dry_mass") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r root dry mass boxplot experiment 2}
exp2_root_dry <- ggplot(exp_2, aes(x=soil, y=roots_dry_mass , fill = soil)) + 
     geom_errorbar(data = summary_N_Root_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "roots_dry_mass") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))
```


```{r aerial dry mass boxplot experiment 2}
exp2_aerial_dry <- ggplot(exp_2, aes(x=soil, y=aerial_dry_mass , fill = soil)) + 
     geom_errorbar(data = summary_N_aerial_dry, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() + labs( x= "Soil Treatment", y= "aerial_dry_mass") + scale_fill_discrete(name = "Legend", labels = c("Naive", "Treated"))+
      theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))

```

```{r combined roots, aerial, total boxplot experiment 2}

long_data_exp2 <- pivot_longer(data = exp_2, 
                          cols = c(total_dry_mass, roots_dry_mass, aerial_dry_mass,percentage_roots, percentage_aerial ), 
                          names_to = "Measurement", 
                          values_to = "Mass")
long_data_exp2$Measurement <- as.factor(long_data_exp2$Measurement)
long_data_exp2$ID <- paste0(long_data_exp2$soil, "_", long_data_exp2$Measurement)
long_exp2 <- long_data_exp2[,-(2:5)]

long_exp2$ID <- paste0(long_data_exp2$soil, "_", long_data_exp2$Measurement)

summary_long_exp2 <- data_summary(data = long_exp2,varname = "Mass",groupnames = c("ID"))

new_cols <- c("soil", "Measurement","dry","mass")
summary_long_exp2 <- separate(summary_long_exp2, col = ID, into = new_cols, sep = "_" )
summary_long_exp2$Measurement <- paste0(summary_long_exp2$Measurement, "_", summary_long_exp2$dry, "_",summary_long_exp2$mass)

dry_mass_allocation_box <- ggplot(long_exp2, aes(x=soil, y=Mass , fill = Measurement)) + 
  labs(x= "Cell Treatment", y= "Root mass") + 
   
  scale_fill_discrete(name = "Legend", labels = c("C_L", "C_R", "EC_L", "EC_R", "MFC_L", "MFC_R", "EV_L", "EV_R"))+
  geom_boxplot(position = position_dodge(0.8)) +
  scale_x_discrete(limits = c("C", "EC", "MFC", "EV"))+
  theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))



ggplot(long_exp2, aes(x=soil,y=Mass, fill = Measurement))+

   geom_errorbar(data = summary_long_exp2, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.8),stat = "identity")+
   
    geom_boxplot(position = position_dodge(0.8))+
    theme(legend.text = element_text(size = 16),
        legend.title = element_text(size = 18),
        axis.title = element_text(size = 14))

```
```{r}
ggplot(long_exp2, aes(x = soil, y = Mass, fill = Measurement)) +
  geom_bar(
    aes(color = Measurement, fill = soil),
    stat = "identity", position = position_stack()
    ) +
  scale_color_manual(values = c("#0073C2FF", "#EFC000FF"))+
  scale_fill_manual(values = c("#0073C2FF", "#EFC000FF"))
```

colours <- c("green", "brown","red", "green", "brown", "red")




dry_mass_allocation_box <- ggplot(long_data_exp2, aes(x = ID, y = mass, fill = ID)) + 
  geom_errorbar(data = summary_long_exp2, aes(ymin = min, ymax = max), width=.2,
                 position=position_dodge(.9),stat = "identity")+
  geom_boxplot() +
  labs(x = "Soil treatmet", y = "Mass") +  
  scale_fill_discrete(name = "Legend", labels = c("Naive Aerial Mass", "Naive Root Mass", "Naive Total Mass", "Treated Aerial", "Treated Root", "Treated Total"))+
   scale_x_discrete(labels = c("Aerial", "Root", "Total", "Aerial", "Root", "Total"))+
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        axis.title = element_text(size = 14))

dry_mass_allocation_box +  scale_fill_manual(values = colours) 

dry_mass_allocation_box +scale_fill_discrete(name = "Legend", labels = c("Naive Aerial Mass", "Naive Root Mass", "Naive Total Mass", "Treated Aerial", "Treated Root", "Treated Total"))

```{r mV and uA line plots dont run}
MFC_mV_line<-ggplot(MFC_mV, aes(x=as.Date(day), y=mV, group=Cell_ID)) +
  geom_line(aes(color=Cell_ID), size =1)+
  geom_point(aes(color=Cell_ID))
MFC_mV_line +theme_minimal()




MFC_uA_line<-ggplot(MFC_uA, aes(x=as.Date(day), y=uA, group=Cell_ID)) +
  geom_line(aes(color=Cell_ID), size =1)+
  geom_point(aes(color=Cell_ID))
MFC_uA_line +theme_minimal()



EV_uA_line<-ggplot(EV_uA, aes(x=as.Date(day), y=uA, group=Cell_ID)) +
  geom_line(aes(color=Cell_ID), size =1)+
  geom_point(aes(color=Cell_ID))
EV_uA_line +theme_minimal()

```

```{r dont run}
calculate_percentage_difference <- function(column1, column2) {
  # Combine columns into a data frame
  data <- data.frame(Column1 = column1, Column2 = column2)
  
  # Convert group column to factor
  data$Column1 <- as.factor(data$Column1)
  
  # Calculate means for each group in column one
  means <- aggregate(. ~ Column1, data, mean)
  
  # Calculate percentage difference between means for each group in column one
  means$Percentage_Difference <- c(NA, means$Column2[-1] / lag(means$Column2[-1]) * 100)
  
  # Output results as a matrix
  result_matrix <- as.matrix(means)
  colnames(result_matrix) <- c("Group", "Mean_Column2", "Percentage_Difference")
  
  return(result_matrix)
}

# Example usage:
column1 <- c("A", "A", "B", "B", "C", "C")
column2 <- c(1, 2, 1, 2, 3, 4)

result <- calculate_percentage_difference(column1, column2)
print(result)

```


```{r dont run}
library(dplyr)

summarize_table <- function(data, group_column) {
  # Group by the specified column and calculate means
  summary_table <- data %>%
    group_by({{ group_column }}) %>%
    summarise(across(where(is.numeric), mean, na.rm = TRUE))
  
  return(summary_table)
}

# Example usage:
# Sample data
set.seed(123)
sample_data <- data.frame(
  Group = rep(letters[1:3], each = 5),
  Numeric1 = rnorm(15),
  Numeric2 = rnorm(15),
  Numeric3 = rnorm(15)
)

# Summarize the table by calculating means for each group in the 'Group' column
leaves.avg<-summarize_table(leaves_no_MFC, treatment)
roots.avg<-summarize_table(roots_no_MFC, treatment)
EV.avg<-summarize_table(EV, LorR)


```